\bf The \texttt{forestecology} R package for modeling interspecies competition between trees

### 2021-02-18 etc/paper/forestecology.tex
Done "Fit model & make predictions" & "Visualize model results" section

\hypertarget{model}{%
\subsection{Model}\label{model}}

Describe model specifics.

@@

  \begin{enumerate}
  \def\labelenumii{\arabic{enumii}.}
  \tightlist
  \item
    Define buffer region trees.
  \item
    Add spatial cross-validation block information.
  \end{enumerate}
\item
  Identify all focal and corresponding competitor trees.
\item
  Fit model and make predictions.
\item
  Additionally: Evaluate model performance using spatial
  cross-validation.
\item
  Additionally: Evaluate the effect of competitor species identity using
  permutation tests.
\end{enumerate}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{library}\NormalTok{(lubridate)}
\KeywordTok{library}\NormalTok{(sf)}
\KeywordTok{library}\NormalTok{(forestecology)}
\KeywordTok{library}\NormalTok{(blockCV)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer), }
               \DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\hypertarget{identify-all-focal-and-corresponding-competitor-trees}{%
\subsection{Identify all focal and corresponding competitor
trees}\label{identify-all-focal-and-corresponding-competitor-trees}}

@@

@@

We estimate the model's parameters using Bayesian linear regression
implemented in the \texttt{fit\_bayesian\_model()} function. TODO:
define all parameters

Using these posterior means, we then use the posterior predictive
distribution to obtain fitted/predicted values \(\widehat{y}\) of the
DBH for each focal tree using the \texttt{predict\_bayesian\_model()}.
These \(\widehat{y}\) can then be compared to the observed \(y\) DBH's
to compute the root mean-square error, a measure of a model's predictive
error which has the same units as the observed data \(y\).

\hypertarget{big-woods}{%
\subsubsection{Big Woods}\label{big-woods}}

For the Michigan Big Woods data we present two use cases of the model
fitting and prediction scheme. The first use case is the simplest where
we assess the fit of the model using root mean squared error. The second
use case then answers the question of whether species competitor
identity matters using permutation test.

For the first use case, we fit the linear model specified in Equation
XXX to our data frame of type \texttt{focal\_vs\_comp}. This
input/outputs of the \texttt{fit\_bayesian\_model()} function are lists
of the prior/posterior means of parameters of the linear regression
specified in XXX. Generally speaking, there are two classes of
regression parameters: \(\beta\) main effects and \(\lambda\)
competitive effects. In the upcoming Section
\ref{viz-posterior-distributions}, we will present code visualizing this
posterior distributions.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{comp_bayes_lm_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

This output of posterior parameters for the specified competition model
are then used along with the posterior predictive distribution encoded
in \texttt{predict\_bayesian\_model()} to return predicted growths for
each individual tree. We join these predicted growths to the original
growth data frame.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_bw, focal_vs_comp_bw))}
\end{Highlighting}
\end{Shaded}

We then use the \texttt{rmse()} function from the \texttt{yardstick}
package to obtain the root mean squared error of the observed versus
fitted values of growth.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\end{Highlighting}
\end{Shaded}

The second use case is near identical to the first, but with a small
change in the code to test whether the identity of the competitor
matters. By adding a \texttt{run\_shuffle\ =\ TRUE} argument to
\texttt{fit\_bayesian\_model()}, for each focal tree its competitor
trees' species identity will be ``shuffled'' randomly much like in a
permutation test. By shuffling these species labels we are effectively
fitting the model under a null model that competitor species identity
does not matter. If the ``shuffled'' RMSE's are consistently lower than
the unshuffled RMSE corresponding to the observed data, then we have
evidence to suggest that competitor identity matters to competitive
interactions.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{comp_bayes_lm_bw_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_bw_shuffle, focal_vs_comp_bw))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat_shuffle) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\end{Highlighting}
\end{Shaded}

The RMSE is fact lower for the non-shuffled version, indicative of a
better model fit. This gives support for the idea that competitor
identity does matter for competitive interactions. In
\citet{allen_permutation_2020} we run this shuffle a large number of
times to construct a full permutation distribution to show that this
difference is robust to resampling variation.

\hypertarget{scbi}{%
\subsubsection{SCBI}\label{scbi}}

In the case of the SCBI data, we once again perform the same model
fitting and computing of fitted growths as with the Big Woods data, but
this time we map the residuals of the observed minus fitted values to
look for spatial patterns.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\end{Highlighting}
\end{Shaded}

In Figures \ref{fig:scbi-model-residuals} and
\ref{fig:scbi-model-residuals-2} we present the residuals.

\hypertarget{viz-posterior-distributions}{%
\subsection{Visualize model results}\label{viz-posterior-distributions}}

Lastly, we return to the model fits from Section \ref{model-fit-predict}
and present tools to visually explore the posterior distributions of all
parameters in our model. There are two main groups of parameters to
consider. The \(\beta\) coefficients tell us about how fast each species
grows and how this depends on DBH while the full matrix of \(\lambda\)
values describe the competitive effects between pairs of species. There
is a rich literature on this matrix (cite).

DO WE NEED TO DESCRIBE MECHANICS? Because of the structure of the
\texttt{bw\_fit\_model} object we cannot simply draw these curves based
on the posterior distribution. \texttt{bw\_fit\_model()} gives the
parameters \emph{compared} to a baseline. This is not of direct
interest. So to display these parameters, as we care about them, we have
to sample from the baseline distribution and from the comparison one to
get the posterior distribution of interest.

\hypertarget{big-woods-1}{%
\subsubsection{Big Woods}\label{big-woods-1}}

Here we re-run the model fit to the Big Woods data from Section
\ref{model-fit-predict}, but this time use ``family'' as the group for
comparison which has. This makes the posterior distributions easier to
follow. Also, surprisingly, grouping by family performed just as well as
grouping by species \citet{allen_permutation_2020}. First we re-run
\texttt{create\_focal\_vs\_comp()} and \texttt{fit\_bayesian\_model()}
with no permutation shuffling with the grouping variable as family.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ family }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_bw, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}

\NormalTok{comp_bayes_lm_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Now the posterior parameter outputs of \texttt{fit\_bayesian\_model()}
are passed to \texttt{plot\_bayesian\_model\_parameters()} to generate
visualizations of the posterior parameters. These visualizations are
displayed in Figure 5 of \citet{allen_permutation_2020}. For simplicity
we only plot a subset of the species families.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"cornaceae"}\NormalTok{, }\StringTok{"fagaceae"}\NormalTok{, }\StringTok{"hamamelidaceae"}\NormalTok{, }\StringTok{"juglandaceae"}\NormalTok{, }
                 \StringTok{"lauraceae"}\NormalTok{, }\StringTok{"rosaceae"}\NormalTok{, }\StringTok{"sapindaceae"}\NormalTok{, }\StringTok{"ulmaceae"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The output is a list with three plots stored. Figure
\ref{fig:bw-posterior-viz-beta0} The element \texttt{beta\_0} gives the
baseline growth intercept \(\beta_0\), i.e., how fast an individual of
each group grows independent of DBH).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{)}
\NormalTok{plot1}
\end{Highlighting}
\end{Shaded}

Figure \ref{fig:bw-posterior-viz-beta-dbh} Next \texttt{beta\_dbh} gives
the slope for DBH slope \(\beta_{dbh,i}\) for each group.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{)}
\NormalTok{plot2}
\end{Highlighting}
\end{Shaded}

Finally Figure \ref{fig:bw-posterior-viz-lambda} \texttt{lambda} gives
the competition coefficients \(\lambda\).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
\NormalTok{plot3}
\end{Highlighting}
\end{Shaded}

\hypertarget{scbi-1}{%
\subsubsection{SCBI}\label{scbi-1}}

We revisit the posterior parameters for the SCBI from Section
\{model-fit-predict\}, but this time only focus on the \(\lambda\)
competition coefficients.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"quru"}\NormalTok{, }\StringTok{"litu"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{, }\StringTok{"cato"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
\NormalTok{plot3}
\end{Highlighting}
\end{Shaded}

Add explanation here.

\hypertarget{spatial-cross-validation}{%
\subsection{Run spatial
cross-validation}\label{spatial-cross-validation}}

@@

\hypertarget{big-woods-2}{%
\subsubsection{Big Woods}\label{big-woods-2}}

Applying this spatially cross-validated model fit yields an RMSE is
higher than that when the model is fit without cross validation. In
other words, our model fits in \ref{model-fit-predict} were overly
optimistic in the model's fitting power, whereas a cross-validated
results yield an estimate that is closer to the truth. See
\citet{allen_permutation_2020} for more discussion of this.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{cv_grid =}\NormalTok{ blocks_bw)}

\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\end{Highlighting}
\end{Shaded}

\hypertarget{scbi-2}{%
\subsubsection{SCBI}\label{scbi-2}}

@@

\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\end{Highlighting}
\end{Shaded}

>>>>

\hypertarget{competition-model}{%
\subsection{Model for growth of tree}\label{competition-model}}

While there are a littany of models one can consider, in PLOSOne we
considered a simple one.

Describe model specifics. Bayesian linear regression model.

Link to \url{https://doi.org/10.1371/journal.pone.0229930.s004}

@@

  \begin{enumerate}
  \def\labelenumii{\arabic{enumii}.}
  \tightlist
  \item
    Define buffer region trees.
  \item
    Add spatial cross-validation block information.
  \end{enumerate}
\item
  Identify all focal and corresponding competitor trees.
\item
  Fit model, make predictions, and visualize model results
\end{enumerate}

Additionally,

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Evaluate the effect of competitor species identity using permutation
  tests.
\item
  Evaluate model performance using spatial cross-validation.
\end{enumerate}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{library}\NormalTok{(lubridate)}
\KeywordTok{library}\NormalTok{(sf)}
\KeywordTok{library}\NormalTok{(forestecology)}
\KeywordTok{library}\NormalTok{(blockCV)}
\KeywordTok{library}\NormalTok{(patchwork)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\hypertarget{focal-vs-comp}{%
\subsection{Identify all focal and corresponding competitor
trees}\label{focal-vs-comp}}

@@

Now that we've identified all focal and corresponding competitor trees
and saved this information in a data frame of type
\texttt{focal\_vs\_comp}, the next step in our analysis sequence is to
fit a model for the growth of all focal trees. Currently the
\texttt{forestecology} package can only fit the competition Bayesian
linear regression model outlined in Section \ref{competition-model}
using the \texttt{comp\_bayes\_lm()} function. However, any model
implemented in a function that similarly takes an input data frame of
type \texttt{focal\_vs\_comp} as an argument can be easily swapped in.
In the case of our specific competition Bayesian linear regression
model, we can also specify prior distributions on all parameters of
interest (here chosen to be the defaults).

The returned \texttt{comp\_bayes\_lm\_scbi} output is an object of S3
class type \texttt{comp\_bayes\_lm} which contains the posterior values
of all parameters in our competition Bayesian linear regression. This
class of object includes generic methods implemented for
\texttt{print()}, \texttt{predict()}, and \texttt{ggplot2::autoplot()}.
First the generic for \texttt{print()} displays the names of all prior
\& posterior parameters (accessible via
\texttt{comp\_bayes\_lm\_scbi\$prior\_params} and
\texttt{comp\_bayes\_lm\_scbi\$post\_params}) along with the model
formula:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{comp_bayes_lm_scbi}
\CommentTok{## Bayesian linear regression model parameters with a multivariate Normal likelihood. See ?comp_bayes_lm for details:}
\CommentTok{## }
\CommentTok{##   parameter_type           prior posterior}
\CommentTok{## 1 Inverse-Gamma on sigma^2 a_0   a_star   }
\CommentTok{## 2 Inverse-Gamma on sigma^2 b_0   b_star   }
\CommentTok{## 3 Multivariate t on beta   mu_0  mu_star  }
\CommentTok{## 4 Multivariate t on beta   V_0   V_star   }
\CommentTok{## }
\CommentTok{## Model formula:}
\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acpl * sp + acru * sp + acsp * sp + aial * sp + amar * sp + astr * sp + beth * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + casp * sp + cato * sp + ceca * sp + ceoc * sp + chvi * sp + coal * sp + coam * sp + cofl * sp + crpr * sp + crsp * sp + divi * sp + elum * sp + eual * sp + fagr * sp + fram * sp + frni * sp + frpe * sp + frsp * sp + havi * sp + ilve * sp + juci * sp + juni * sp + juvi * sp + libe * sp + litu * sp + loma * sp + nysy * sp + pato * sp + pipu * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prpe * sp + prse * sp + prsp * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru * sp + qusp * sp + quve * sp + rhpe * sp + romu * sp + rops * sp + rual * sp + rupe * sp + ruph * sp + saal * sp + saca * sp + tiam * sp + ulam * sp + ulru * sp + ulsp * sp + unk * sp + viac * sp + vipr * sp + vire * sp}
\end{Highlighting}
\end{Shaded}

Next, the generic for \texttt{predict()} takes as inputs the posterior
parameter values in \texttt{comp\_bayes\_lm\_scbi} and the predictor
variables in \texttt{newdata} and outputs a vector of fitted/predicted
values \(\widehat{y}\) of the DBH for each focal tree (as returned by
the posterior predictive distribution).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi}
\CommentTok{## # A tibble: 6,296 x 8}
\CommentTok{##    focal_ID focal_sp   dbh foldID                  geometry  growth}
\CommentTok{##       <dbl> <fct>    <dbl> <fct>                    <POINT>   <dbl>}
\CommentTok{##  1        4 nysy     13.6  1                   (14.2 428.5)  0.103 }
\CommentTok{##  2        5 havi      8.8  1                    (9.4 436.4)  0.150 }
\CommentTok{##  3       79 tiam     47.7  1                     (40 381.1) -0.161 }
\CommentTok{##  4       80 caca      5.15 1                   (38.7 421.7)  0.253 }
\CommentTok{##  5       96 libe      2.3  1                       (60 310)  0.262 }
\CommentTok{##  6      101 litu     65.4  1                   (47.1 567.3)  0.552 }
\CommentTok{##  7      102 astr      1.99 1                   (40.8 575.5)  0.0954}
\CommentTok{##  8      126 cato     37.4  1                   (60.6 400.2)  0.165 }
\CommentTok{##  9      127 caca      8.72 1                   (72.7 514.1)  0.0370}
\CommentTok{## 10      139 astr      1.71 1                   (96.7 315.1)  0.0549}
\CommentTok{## # ... with 6,286 more rows, and 2 more variables: comp <list>,}
\CommentTok{## #   growth_hat <dbl>}
\end{Highlighting}
\end{Shaded}

We can then compare the observed and fitted/predicted growths to compute
the root mean squared error of our model fit.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{model_rmse <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\NormalTok{model_rmse}
\CommentTok{## [1] 0.1281398}
\end{Highlighting}
\end{Shaded}

Lastly, the generic for \texttt{ggplot2::autoplot()} allows us to

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Plot posteriors for only a subset of species}
\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"quru"}\NormalTok{, }\StringTok{"litu"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{, }\StringTok{"cato"}\NormalTok{)}

\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}

\CommentTok{# Combine plots using patchwork}
\NormalTok{(plot1 }\OperatorTok{|}\StringTok{ }\NormalTok{plot2) }\OperatorTok{/}\StringTok{ }\NormalTok{plot3}
\end{Highlighting}
\end{Shaded}

{\centering \includegraphics[width=1\linewidth]{Figures/scbi-posterior-viz-1}

\caption{Posterior distributions of all parameters for a subset of species.}\label{fig:scbi-posterior-viz}
\end{figure}

\hypertarget{evaluate-the-effect-of-competitor-species-identity-using-permutation-tests}{%
\subsection{Evaluate the effect of competitor species identity using
permutation
tests}\label{evaluate-the-effect-of-competitor-species-identity-using-permutation-tests}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{)}

\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi)}
\NormalTok{  )}

\NormalTok{model_rmse <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\NormalTok{model_rmse}
\end{Highlighting}
\end{Shaded}

\hypertarget{spatial-cross-validation}{%
\subsection{Evaluate model performance using spatial
cross-validation}\label{spatial-cross-validation}}

@@

@@

\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\CommentTok{## [1] 0.1446232}
\end{Highlighting}
\end{Shaded}
